---
title: "Mushroom Classification"
author: "Sanjay Lokula"
date: "24/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading the Necessary Libraries

```{r}
library(stringr)
library(stringi)
library(car)
library(dplyr)
library(ggplot2)
library(dplyr)
library(corrplot)
library(Seurat)
library(caret)
library(plyr)
```


#### Loading the Dataset

```{r}
mushroom_data <- read.csv(file.choose(), stringsAsFactors = FALSE)
```


#### Summary of the dataset

```{r}
summary(mushroom_data)
```


#### Class identification

```{r}
str(mushroom_data)
# All type of data are factor

```


#### Missing value analysis
# Let's see is there any NA value and is there any column needs a better name
```{r}
sapply(mushroom_data,function(x)sum(is.na(x)))
# There is no missing value in any column, and all column's name are clear

```

### EDA (Exploratory Data Analysis)

#### Feature Analysis

```{r}

class <- plyr::count(mushroom_data$class)
print(sprintf("Edible: %d | Poisonous: %d | Percent of poisonous classes: %.1f%%",class$freq[1],class$freq[2], round(class$freq[1]/nrow(mushroom_data)*100,1)))
```


#### Analyzing the attributes with low numbers of categories
#### Analyzing "bruises"

```{r}
ggplot(mushroom_data, aes(bruises, fill = class)) + geom_bar(position = "dodge")
```

#### Analyzing gill.attachment
```{r}
ggplot(mushroom_data, aes(gill.attachment, fill = class)) + geom_bar(position = "dodge")

```



#### gill.spacing
```{r}
ggplot(mushroom_data, aes(gill.spacing, fill = class)) + geom_bar(position = "dodge")

```


#### Almost equally distributed classes 

#### gill.size
```{r}
ggplot(mushroom_data, aes(gill.size, fill = class)) + geom_bar(position = "dodge")

```



#### stalk.shape
```{r}
ggplot(mushroom_data, aes(stalk.shape, fill = class)) + geom_bar(position = "dodge")

```


#### odor
```{r}
ggplot(mushroom_data, aes(odor, fill = class)) + geom_bar(position = "dodge")
#Fully Edible: 'a' & 'l'
#Mostly Edible: 'n' (A very small portion of this class fall under poisonous)
#Fully poisonous: 'c', 'f', 'm','p', 's' & 'y'
```




#### Class analysis
##### Calculate number of class for each variable
```{r}
num_of_class <-cbind.data.frame(Var=names(mushroom_data), Total_Class=sapply(mushroom_data,function(x){as.numeric(length(levels(factor(x))))}))
num_of_class

```

### Features analysis Part 2
#### In case if needs to do further investigate so transfter all factor to numeric

```{r}
mushroom_numeric<- sapply(mushroom_data[,1:23], function (x) as.numeric(as.factor(x)))
head(mushroom_numeric)
```




### Correlation analysis
#####  Delete no.17 column of veil-type since it has only one type that is partial
```{r}
cor_mushroom <- cor(mushroom_numeric[, -17])
corrplot.mixed(cor_mushroom,order="AOE")


CombinePlots(plots=list(a,b,c,d,e,f))
a= ggplot(mushroom_data) + geom_bar(aes(x = class))+
  ggtitle("Distribution of class") + ylim(0, 5000)

b = ggplot(mushroom_data) + geom_bar(aes(x = cap.shape))+
  ggtitle("Distribution of cap.shape") + ylim(0, 4000)

c = ggplot(mushroom_data) + geom_bar(aes(x = cap.surface))+
  ggtitle("Distribution of cap.surface") + ylim(0, 4000)

d = ggplot(mushroom_data) + geom_bar(aes(x = cap.color))+
  ggtitle("Distribution of cap.color") + ylim(0, 3000)

e = ggplot(mushroom_data) + geom_bar(aes(x = bruises))+
  ggtitle("Distribution of bruises") + ylim(0, 5000)

f = ggplot(mushroom_data) + geom_bar(aes(x = odor))+
  ggtitle("Distribution of odor") + ylim(0, 4000)

```

